Pillar 1 - Sequential Pattern Mining

### 2019-02-12 Pillar 1 - Sequential Pattern Mining.Rmd
corrects confidence in TPF sequences

# Type level
statistics_type <- d_transcript %>%
  inner_join(d_course, by = "Course ID") %>%
  filter(!is.na(Type)) %>%
  group_by(Type) %>%
  provide_statistics()

@@

#rm(provide_statistics, statistics_student, statistics_course, statistics_cluster,
#     statistics_concentration, statistics_year, statistics_level)
```

@@

### Transcript with Preceding courses per student
```{r}

#
# List of all courses
course_all <- unique(d_transactions$taken_PF_HL$item)

#
# courses taken so far and not taken so far
d_cumulative_transcript <- d_transactions$taken_PF_HL %>%

  group_by(
    sequenceID,
    eventID
    ) %>%
  summarize(
    course_current = list(
      unique(item)
      )
    ) %>%

  group_by(
    sequenceID
    ) %>%
  mutate(
    course_past = Reduce(
      f = paste, 
      x = course_current, 
      accumulate = TRUE
      ) %>%
      lag
    ) %>%

  rowwise %>% # equivalent to "group by row"
  mutate(
    course_so_far     = list(union  (course_past, course_current)),
    course_not_so_far = list(setdiff(course_all , course_so_far ))
    ) %>%
  ungroup

rm(course_all)
```

@@

SR_TPF are manually computed
```{r}
# current, past, not_yet
SR_TPF <- d_cumulative_transcript                  %>%
  select(sequenceID, eventID, item, course_not)    %>%
  unnest(item, .drop = F)                          %>%
  unnest(course_not)                               %>%
  left_join(
    select(d_transactions, 
           sequenceID, eventID, item, PF), 
    by= c("sequenceID", "eventID", "item")
    )                                              %>%
  select(sequenceID,eventID, course_not, item, PF) %>%
  rename(
    lhs="course_not", 
    rhs.x = "item", 
    rhs.PF = "PF")                                 %>%
  mutate(num_transactions = n())                   %>%
  group_by(lhs)                                    %>%
  mutate(
    num_lhs = n(),
    support_lhs         = num_lhs / num_transactions
    )                                              %>%
  group_by(rhs.x)                                  %>%
  mutate(
    num_rhs.x           = n(),
    support_rhs.x       = num_rhs.x / num_transactions
         )                                         %>%
  group_by(rhs.x,rhs.PF)                           %>%
  mutate(
    num_rhs.PF          = n(),
    support_rhs.PF      = num_rhs.PF / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x)                             %>%
  mutate(
    num_rule_x          = n(),
    support_rule_rhs.x  = num_rule_x / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x, rhs.PF)                     %>%
  mutate(
    num_rule_PF          = n(),
    support_rule_rhs.PF  = num_rule_PF / num_transactions,
    support_rhs_bounded  = num_rule_PF / num_rule_x
         )                                         %>%
  ungroup()                                        %>%
  mutate(
    confidence                   = support_rule_rhs.PF / support_lhs,
    confidence_bounded           = support_rhs_bounded / support_lhs, #not sure about this 
    lift                         = confidence / support_rhs.PF,
    lift_bounded                 = confidence_bounded /support_rhs_bounded #not sure about this
  )
```

>>>>

@@

rm(provide_statistics, statistics_student, statistics_course, statistics_cluster,
     statistics_concentration, statistics_year, statistics_level)
```

@@

### Transcript with Preceding courses per student
```{r}
d_cumulative_transcript <-  d_transactions        %>%
  select(sequenceID, item, eventID)  %>%
  arrange(sequenceID, eventID, item) %>%
  group_by(sequenceID, eventID)      %>%
  summarize_all(paste, collapse=" ") %>%
  arrange(sequenceID, eventID)       %>%
  mutate(
    course_all  = paste(unique(d_transactions$item), collapse = " "),
    course_all  = str_split(course_all, pattern = " "),

    course_now  = str_split(item, pattern = " "),

    course_past     = Reduce(paste, item, accumulate=T),
    course_past     = lag(course_past),
    course_past     = str_split(course_past, pattern = " ")) %>%
  ungroup() %>%
    mutate(
    course_past_now = purrr::map2(.x = course_past,
                                      .y = course_now,
                                      .f = union),
    course_not = purrr::map2( .x = course_all,
                              .y = course_past_now,
                              .f = setdiff)
    ) %>%
  mutate(item = str_split(item, pattern = " "))
```

@@

SR_TPF are manually computed
```{r}
# current, past, not_yet
SR_TPF <- d_cumulative_transcript                  %>%
  select(sequenceID, eventID, item, course_not)    %>%
  unnest(item, .drop = F)                          %>%
  unnest(course_not)                               %>%
  left_join(
    select(d_transactions, 
           sequenceID, eventID, item, PF), 
    by= c("sequenceID", "eventID", "item")
    )                                              %>%
  select(sequenceID,eventID, course_not, item, PF) %>%
  rename(
    lhs="course_not", 
    rhs.x = "item", 
    rhs.PF = "PF")                                 %>%
  mutate(num_transactions = n())                   %>%
  group_by(lhs)                                    %>%
  mutate(
    num_lhs = n(),
    support_lhs         = num_lhs / num_transactions
    )                                              %>%
  group_by(rhs.x)                                  %>%
  mutate(
    num_rhs.x           = n(),
    support_rhs.x       = num_rhs.x / num_transactions
         )                                         %>%
  group_by(rhs.x,rhs.PF)                           %>%
  mutate(
    num_rhs.PF          = n(),
    support_rhs.PF      = num_rhs.PF / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x)                             %>%
  mutate(
    num_rule_x          = n(),
    support_rule_rhs.x  = num_rule_x / num_transactions
  )                                                %>%
  group_by(lhs, rhs.x, rhs.PF)                     %>%
  mutate(
    num_rule_PF          = n(),
    support_rule_rhs.PF  = num_rule_PF / num_transactions,
    support_rhs_bounded  = num_rule_PF / num_rule_x
         )                                         %>%
  ungroup()                                        %>%
  mutate(
    confidence                   = support_rule_rhs.PF / support_lhs,
    confidence_bounded           = num_rule_PF/num_rule_x, #support_rhs_bounded / support_lhs, #not sure about this 
    lift                         = confidence / support_rhs.PF,
    lift_bounded                 = confidence_bounded /support_rhs_bounded #not sure about this
  )
```
